# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r}
unzip('activity.zip')
origin <- read.csv('./activity.csv')
```


## What is mean total number of steps taken per day?

```{r}
data <- origin[!is.na(origin[, 1]), ]
s <- split(data, data$date)
a <- sapply(s, function(x)sum(x[, 1]))
hist(a, main="Histogram of total steps per day", breaks = 15, xlab = "total number of steps per day")
```

```{r stats, echo=FALSE}
mean0 <- mean(a)
median0 <- median(a)
```
The mean of the total number of steps taken per day is `r floor(mean0)`;  
the median of the total number of steps taken per day is `r median0`.

## What is the average daily activity pattern?
```{r}
s <- split(origin, origin$interval)
a <- sapply(s, function(x)mean(x[, 1], na.rm = TRUE))
x <- 5 * (1:length(a))
plot( x, a, type = "l", xlab = "time series", ylab = "the average number of steps", main = "Activity Pattern")
```
```{r, echo=FALSE}
m <- which.max(a)
```
`r m` index of 5-minute interval, on average across all days in the dataset, contains the maximum number of steps.


## Imputing missing values
```{r, echo=FALSE}
m1 <- sum(is.na(origin[,1]))
```
The total number of missing values in the dataset is `r m1`.

The following code is to fill in all of the missing step values with the mean steps for that 5-minute interval.
```{r, cache=TRUE}
temp <- sapply(s, function(x)mean(x[, 1], na.rm=TRUE))
for (i in seq(nrow(origin))){
        if (is.na(origin[i, 1])){
                if(!(i%%288==0)){
                        origin[i, 1] = floor(temp[i%%288])
                }else {
                        origin[i, 1] = floor(temp[288])
                }
        }
}
```
```{r}
s1 <- split(origin, origin$date)
a1 <- sapply(s1, function(x)sum(x[, 1]))
hist(a1, main="Histogram of total steps per day after filling missing data", breaks = 15, xlab = "total number of steps per day")
```
```{r, echo=FALSE}
mean1 <- mean(a1)
median1 <- median(a1)
```
After filling the missing data, 
The mean of the total number of steps taken per day is `r floor(mean1)`;  
the median of the total number of steps taken per day is `r floor(median1)`.  
These values differ from the estimates from the first part of the assignment.  
Imputing missing data will increase on the estimates of the total daily number of steps. 


## Are there differences in activity patterns between weekdays and weekends?
```{r}
days <- as.Date(origin[, 2])
wd <- weekdays(days) == "Sunday" | weekdays(days) == "Saturday"
wd[wd == TRUE ] = "weekend"
wd[wd == FALSE] = "weekday"
wd <- factor(wd)
origin <- cbind(origin, wd)

library("reshape2")
library("lattice")
Melt <- melt(origin, id=c("interval","wd"), measure.vars="steps")
Cast <- dcast(Melt, interval + wd ~ variable, mean)
#head(Cast)
xyplot(steps ~ interval | wd, data = Cast, type="l", ylab="Number of steps", layout = c(1,2))

```


